### 2023-01-31 outputs/paper.qmd
draft

---
title: "doupi3"
subtitle: "sid"
author: "Mengze Zhang"
thanks: 'Code and data are available at [github.com/MissyZhang/Crime_Data_by_Neighbourhood](https://github.com/MissyZhang/Crime_Data_by_Neighbourhood).'
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "we consider data about and find that"
format: pdf
bibliography: reference.bib
---

```{r}
#| include: FALSE
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palettetown)
library(here)
library(ggplot2)
library(knitr)
library(kableExtra)
library(patchwork)
```

# Introfuction

# Data

## Data Source ##

## Data Collection ##

## Data Analysis ##

Analysis for this project uses the R statistical programming language [@citeR], and more specifically, the `tidyverse` package for data manipulation [@citetidyverse].Because the data is managed using R Projects, `here` is used to reference file locations [@citehere].

```{r}
#| include: FALSE
# Import the data that was first imported by 00-download_data.R
# and then cleaned by 01_data_cleaning.R
cleaned_data <- readr::read_csv(here::here("inputs/data/cleaned_data.csv"))

# Classify neighborhoods by population
cleaned_data$neighbothood_size <- cut(cleaned_data$population2021, breaks = c(-Inf,10000, 15000, 25000,Inf), labels = c("Small", "Medium", "Large", "Giant"), right = FALSE)

```

```{r}
#| echo: FALSE
#| warning: False
#| message: false
crime_rate_by_nbsize2021 <- cleaned_data |>
  select(neighbothood_size, 
         assault_rate2021,  
         auto_theft_rate2021, 
         break_and_enter_rate2021,
         robbery_rate2021,
         theft_over_rate2021,
         homicide_rate2021,
         shootings_rate2021,
         theftfrom_motor_vehicle_rate2021
         ) |>
  group_by(neighbothood_size)|>
  summarise(mean_assault_rate=mean(assault_rate2021, na.rm = TRUE),
            mean_auto_theft_rate=mean(auto_theft_rate2021, na.rm = TRUE),
            mean_break_and_enter_rate=mean(break_and_enter_rate2021, na.rm = TRUE),
            mean_robbery_rate=mean(robbery_rate2021, na.rm = TRUE),
            mean_theft_over_rate=mean(theft_over_rate2021, na.rm = TRUE),
            mean_homicide_rate=mean(homicide_rate2021, na.rm = TRUE),
            mean_shootings_rate=mean(shootings_rate2021, na.rm = TRUE),
            mean_theftfrom_motor_vehicle_rate=mean(theftfrom_motor_vehicle_rate2021, na.rm = TRUE)) |>

# Tidy up names so they are better for the table display:
  rename("Assault" = mean_assault_rate,
         "Auto Theft" = mean_auto_theft_rate,
         "Break & Enter" = mean_break_and_enter_rate,
         "Robbery" = mean_robbery_rate,
         "Theft" = mean_theft_over_rate,
         "Homicide" = mean_homicide_rate,
         "Shooting" = mean_shootings_rate,
         "Motor Theft" = mean_theftfrom_motor_vehicle_rate,
         "Neighborhood Size" = neighbothood_size) |>
  mutate_if(is.numeric, round, 2) # round rate averages to 2 decimal places

# Put the results in a table:
pivottable <-as_tibble(cbind(nms = names(crime_rate_by_nbsize2021), t(crime_rate_by_nbsize2021)))

kable(pivottable [-1,],
      col.names = c("Neighborhood Size",
                    "Small",
                    "Medium",
                    "Large",
                    "Giant"),
  caption = "Mean crime rate by neighborhood size in 2021",
      booktabs = TRUE, 
      longtable = TRUE
      ) |>
   kable_styling(bootstrap_options = c("hover", "condensed")) |>

# Style the table
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "3em") %>%
  column_spec(3, width = "3em") %>%
  column_spec(4, width = "3em") %>%
  column_spec(5, width = "3em") %>%
  pack_rows("Crime Type", 1, 8) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))

```
```{r}
#| echo: false
#| message: false
#| warning: false
#| n-col: 2
#homicide(NA), shooting(NA), motor theft
selected_crime_rate2021 <- cleaned_data |>
  select(population2021, homicide_rate2021, shootings_rate2021, theftfrom_motor_vehicle_rate2021)

homicide <-
  selected_crime_rate2021 |>
  ggplot(mapping = aes(x = population2021, y= homicide_rate2021))+
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  labs(
    x = "Population",
    y = "Homocide Rate",
    )

shootings <-
  selected_crime_rate2021 |>
  ggplot(mapping = aes(x = population2021, y= shootings_rate2021))+
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  labs(
    x = "Population",
    y = "Shootings Rate",
    )

motor <-
  selected_crime_rate2021 |>
  ggplot(mapping = aes(x = population2021, y= theftfrom_motor_vehicle_rate2021))+
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  labs(
    x = "Population",
    y = "Motor theft Rate",
    )

(homicide + shootings) /
  motor

```
```{r}
#| include: false
#| echo: false
#| message: false
#| warning: false
mean_crime_rate_each_year <- cleaned_data |>
  select(-population2021,-neighbothood_size) |>
  summarise_all(mean, na.rm = TRUE)

mean_crime_rate_by_year_and_crime_type <- data.frame(
  crime_type = c(rep(x="Assault", 5),
                 rep(x="Auto Theft", 5), 
                 rep(x="Break & Enter", 5),
                 rep(x="Robbery", 5),
                 rep(x="Theft", 5), 
                 rep(x="Homicide", 5),
                 rep(x="Shooting" , 5),
                 rep(x="Motor Theft", 5)
                 ),
  year= c(2017:2021),
  mean_crime_rate = gather(mean_crime_rate_each_year)
)

```

```{r}
#| fig.cap: "Crime rate in Toronto by category and year"
#| fig.width : 8
#| fig.height : 4
#| echo : FALSE
#| message : FALSE
#| warning : FALSE
#| out.width : '80%'
#| fig.align : "center"

mean_crime_rate_by_year_and_crime_type |>
  select(crime_type, year, mean_crime_rate.value) |>
  group_by(crime_type) |>
  ggplot() +
  geom_point(aes(x = year, y = mean_crime_rate.value, color = crime_type), size = 2) +
  scale_color_poke(pokemon = 'Charizard', spread = NULL)+ 
  theme_light() + 
   labs(x = "Year", y = "Mean Crime Rate",
       title = " Assaults Are the Most Frequently Committed Category Offense") + 
  guides(color = guide_legend(title = "Category",override.aes = list(size=3))) +
    theme(plot.title = element_text(face = "bold"))

```
```